---
title: "HW00: Mock submission"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* **Statistics 159/259, Fall 2026**
* Prof. Gaston Sanchez, Department of Statistics, UC Berkeley
* GSIs: Sequoia Andrade & Larissa Arreola Moreno

* This assignment is worth a maximum of **40 points**.
* Assignment type: **individual**.

**Deliverable:** for this assignment, your github repository should contain:

- This notebook, completed so it includes code for plots and simulations, along with your written responses and discussion. Please remember to use markdown headings for each section/subsection so the entire document is readable.

**Collaboration Policy**

Data science is a collaborative activity. While you may talk with others about
the project, we ask that you **write your solutions individually**. If you do
discuss the assignments with others please **include their names** below:

## TODO  - HW00
### Notebook: coding exercise
**Now switch to this notebook** — the rest of Part 2 happens in the cells below, as ordinary R.

This is `hw00` **Q1: dictionaries for counting words**. A common text-processing task: count how often each word appears, using a dictionary.

Fill in `word_freq` below: split `text` into words on whitespace, and build a dictionary mapping each word to how many times it appears.

```{r word-freq}
word_freq <- function(text) {
  # Return a dictionary of word frequencies for the given text.
  freqs <- list()
  # TODO: split text into words, and count them into freqs
  freqs
}


print_vk <- function(pairs) {
  # pairs is a data frame with columns `word`, `count`, already sorted.
  for (i in seq_len(nrow(pairs))) {
    cat(pairs$word[i], ": ", pairs$count[i], "\n", sep = "")
  }
}


freq_summ <- function(freqs, n = 10) {
  # Print a simple summary of a word frequencies dictionary.
  df <- data.frame(word = names(freqs), count = unlist(freqs))
  df <- df[order(-df$count), ]
  cat("Top", n, "words:\n")
  print_vk(head(df, n))
  cat("\nBottom", n, "words:\n")
  print_vk(tail(df, n))
}
```

Try it on a short sample text:

```{r word-freq-check}
sample_text <- "
the quick brown fox jumps over the lazy dog
the dog barks at the fox but the fox runs away
the quick fox is quicker than the lazy dog every single time
"

freqs <- word_freq(sample_text)
freq_summ(freqs, n = 3)
```

## TODO
### Terminal: save your final progress
**Back in the terminal**:

```bash
git clone <link to your repo>
git add HW00.Rmd
git commit -m "Complete word frequency counting exercise"
git log --oneline
```

# Congratulations! You have completed Homework 0!
